\bf On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2024-11-23 Manuscript/manuscript.tex
half way through data section

There are two types of data that are explored throughout this paper. For the
rank-based comparisons described in Section~\ref{sec:rank}, the data described in 
Section~\ref{sec:databeyond} come from male 
athletes who competed in the 110 meter hurdles and \eds{or?} 100 meter dash, and 
female athletes who competed 100 meter hurdles and \eds{or?} 100 meter dash.
Athletes must have competed in the 2022 World Championships and another competition 
(2022 national championships, 2019 World Championships, or 2023 World Championships) 
in order to be included in the rank-based comparisons. 
For the GAMLSS analysis described in Section~\ref{sec:gamlss}, the data described 
in Section~\ref{sec:dataworld} come
from every world championships from 1999 to 2023 and includes reaction times from
the 110 meter hurdles and 100 meter dash. This investigation started 
shortly after the 2022 World Championships, but as the 2023 data has now become 
available, we are pleased to include it in both of our analyses, and it did not
significantly change our results \citep{WAData}. \eds{Should we remove this 
last sentence, since now likely 2024 data is available?}

@@

The motivation for looking beyond World Championships data came from an 
exploratory analysis examining how United States athletes reacted at the 
2022 United States Track and Field Championships.

\subsubsection{2022 National Competitions}\label{sec:datanational}
From June 23-26, 2022, the United States held its Track and Field Championships 
at Hayward Field in Eugene, Oregon to decide who to send to the World 
Championships being held at the same venue in August. Thus, we can establish a 
baseline for the four United States athletes who competed in at least one round 
of both events: Trey Cunningham, Daniel Roberts, Grant Holloway, and Devon Allen.  
Every athlete reacted faster in all the World Track and Field Championships 
races compared to the United States Track and Field Championships races. The 
dataset of the four United States athletes is too small to perform any 
meaningful statistical analysis, so the data for the rank-based analyses was 
expanded to include reaction 
times from more athletes who competed at other national competitions (many of which were 
held from May-July of 2022). This data was difficult to find, as the data is not
centrally located but instead found on different national websites (often 
recorded in the national language without translation).

Each athlete was designated as a ``cluster'' and the stage of competition (national
or international) was specified for each competition within the cluster (athlete).
Thus, this is subunit level grouping where we assign the treatment, national or
international, within each cluster. Cluster sizes range from three to six, with
one cluster of size three, six clusters of size four, two clusters of size five,
and two clusters of size six.
\of{This explanation is going to need to be re-worked now that we have more data.
Should we make a table of some sort? It's going to be very wordy without one}
\eds{We are only doing two analyses with this data, right, one for women 
and one for men?  So perhaps only two sentences needed here? But then 4 sentences
in the next subsection?  Maybe you can just report ranges of cluster sizes and 
median cluster size?}

We chose to include reaction times from preliminary heats, in addition to 
reaction times from semifinals and finals, in this section and in
Section~\ref{sec:data2019}, as removing them would have resulted in a much
narrower analysis with fewer athletes and clusters.

\subsubsection{Other World Track and Field Championships}\label{sec:data2019}
The 2019 and 2023 World Track and Field Championships data is included in the 
larger dataset described in Section~\ref{sec:dataworld}, but we can also use it
similarly to how we used national competitions in Section~\ref{sec:datanational}. 
We can compare the reaction times of athletes who participated in the 2022 World 
Championships with their times at another World Championships: 2019 and/or 2023.

While it is theoretically possible that athletes got faster 
between 2019 and 2022 or 2022 and 2023, it is important to 
remember that these are some of the fastest sprinters on the planet, and thus
improving on elite times is not easy. Once again we repeat a similar procedure
as in Section~\ref{sec:datanational}; we assign each athlete to be a cluster and
within each cluster label times to be from either the 2019, 2022, or 2023 World
Championships. When comparing 2019 and 2022 we look at data from 15
athletes, with two clusters of size two, three clusters of size three, five 
clusters of size four, two clusters of size five, and three clusters of size six.
When comparing 2022 and 2023 we look at data from 22 athletes, with seven 
clusters of size two, three clusters of size three, six clusters of size four,
and six clusters of size five.
\of{This explanation might need to change}

>>>>

This study focuses on reaction times from sprint events where reaction
time plays a critical role. Specifically, we included data from men’s
100-meter dashes and 110-meter hurdles, as well as women’s 100-meter
dashes and 100-meter hurdles. Statistical tests showed no significant
differences in reaction times between these events, supporting their
inclusion in a unified analysis. However, we excluded data from
200-meter dashes and longer events, as their reaction times were found
to be significantly different. This difference is not unexpected, as
reaction time is generally less critical in longer sprint events
compared to shorter distances where explosive starts are paramount.

Two datasets were used to address the study's objectives. To
investigate whether reaction times at the 2022 World Championships
were significantly different from other competitions (Objective 1), we
used data from male athletes in the 110-meter hurdles and 100-meter
dash, and female athletes in the 100-meter hurdles and 100-meter dash,
provided they competed in the 2022 World Championships and at least
one other competition (2022 national championships, 2019 World
Championships, or 2023 World Championships). To assess the
appropriateness of the 0.1-second reaction time threshold (Objective
2), we used reaction times from the 110-meter hurdles and 100-meter
dash across all World Championships held from 1999 to 2023. This
investigation began shortly after the 2022 World Championships, and we
are pleased to include data from 2023, which did not significantly
alter our results \citep{WAData}.

@@

The motivation for including data beyond the World Championships arose from 
an exploratory analysis of how United States athletes performed at the 2022 
United States Track and Field Championships, held from June 23--26, 2022, at 
Hayward Field in Eugene, Oregon. Since the same venue hosted the 2022 World 
Championships in August, this provided a unique opportunity to establish a 
baseline for reaction times.

\subsubsection{2022 National Competitions}
\label{sec:datanational}

To compare reaction times at the 2022 World Championships with those
recorded at national competitions, we constructed a dataset of athletes 
who competed in both settings. Initially, we examined data from four 
United States athletes---Trey Cunningham, Daniel Roberts, Grant 
Holloway, and Devon Allen---who participated in both the United States 
Track and Field Championships and the World Championships. Each athlete 
recorded faster reaction times in all World Championships races compared 
to their performances at the national-level event. However, this small 
sample size was insufficient for meaningful statistical analysis. To 
expand the dataset, we included reaction times from additional national 
competitions held between May and July 2022 across various countries. 
Compiling this data was challenging, as reaction times were not centrally 
located and often required searching country-specific websites, with many 
results recorded in native languages.

\jy{Owen, does this include both dash and hurdle? Could you do a
  pooled analysis since there is no gender difference and event
  differece? Need to summarize how many athletes are included in each
  analysis.}

The final dataset consisted of reaction times from athletes who competed 
in both national and international competitions. Reaction times from 
preliminary heats, semifinals, and finals were included to ensure 
sufficient data for analysis. Excluding preliminary heats would have 
significantly reduced the number of athletes and clusters, narrowing the 
scope of the analysis. Each athlete can be considered as a cluster, with
reaction times from the `treatment' group (2022 World Championships) and 
the `control' group (national competition) within the cluster. Cluster 
sizes ranged from three to six, with a median size of four. This
setting forms the basis for rank-based comparison where the
within-cluster dependence needs to be accounted for.

\of{This explanation is going to need to be re-worked now that we have more data.
Should we make a table of some sort? It's going to be very wordy without one}
\eds{We are only doing two analyses with this data, right, one for women 
and one for men?  So perhaps only two sentences needed here? But then 4 sentences
in the next subsection?  Maybe you can just report ranges of cluster sizes and 
median cluster size?}

\subsubsection{2019 and 2023 World Championships}
\label{sec:data2019}
The 2019 and 2023 World Track and Field Championships data is included in the 
larger dataset described in Section~\ref{sec:dataworld}, but we can also use it
similarly to how we used national competitions in Section~\ref{sec:datanational}. 
We can compare the reaction times of athletes who participated in the 2022 World 
Championships with their times at another World Championships: 2019 and/or 2023.

While it is theoretically possible that athletes got faster 
between 2019 and 2022 or 2022 and 2023, it is important to 
remember that these are some of the fastest sprinters on the planet, and thus
improving on elite times is not easy. Once again we repeat a similar procedure
as in Section~\ref{sec:datanational}; we assign each athlete to be a cluster and
within each cluster label times to be from either the 2019, 2022, or 2023 World
Championships. When comparing 2019 and 2022 we look at data from 15
athletes, with two clusters of size two, three clusters of size three, five 
clusters of size four, two clusters of size five, and three clusters of size six.
When comparing 2022 and 2023 we look at data from 22 athletes, with seven 
clusters of size two, three clusters of size three, six clusters of size four,
and six clusters of size five.

### 2024-11-23 Manuscript/reply.tex
half way through data section

In the revision, we have added explanation about why 100-meter dash
data for both men and wemen were included, while 200-meter dash data
were not included in our analysis in Section~???, paragraph~???.
The subsequent analyses in the manuscript have been updated with the
expanded dataset that included both men and women's 100-meter dash
data; see updated results in Sections~???.
\eds{let's revisit this again after objective 2 analyses are set}

>>>>

In the revision, we have added explanation about why 100-meter dash
data for both men and wemen were included, while 200-meter dash data
were not included in our analysis in the opening paragraph of Section~2.
The subsequent analyses in the manuscript have been updated with the
expanded datasets.
\eds{let's revisit this again after objective 2 analyses are set}
